### 2013-05-01 inst/examples/BUGS/allen.Rmd
Example using custom GP routine.
In this case, both MLE methods fail, while the GP does fine and the parametric Bayesian method performs best.

```{r}
f <- RickerAllee
p <- c(.2, 7.5, 5)
K <- p[2]
allee <- p[3]
```

@@

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 20  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
seed_i <- 111
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 35
```

@@

```{r gp-priors}
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
tau2.p <- c(5,1)
d.p = c(10, 1/0.1, 10, 1/0.1)
nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)
beta = c(0)
priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
```

@@

```{r}
gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, zcov=TRUE, verb=0,
          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
```

@@

```{r, eval=FALSE}
hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")

# Traces 
ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

## A list of functions has been defined in "priors", 
## Here we just make sure to compute priors over the range of values returned by the posteriors
prior_curves <- ddply(posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
})

# Posteriors (easier to read than histograms)
ggplot(posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)

@@

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 0.01, 20)
logr_prior <- function(x) dunif(x, -5, 5)
logtheta_prior <- function(x) dunif(x, -5, 5)
stdQ_prior <- function(x) dunif(x, 0.001, 100)
stdR_prior <- function(x) dunif(x, 0.001, 100)

@@

```{r}
# Summarize the GP model
tgp_dat <- 
    data.frame(  x = gp$XX[[1]], 
                 y = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))

@@

plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

@@

```{r}
MaxT = 1000
matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid)
opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

>>>>

```{r}
f <- RickerAllee
p <- c(2, 8, 5)
K <- p[2]
allee <- p[3]
```

@@

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 30  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
seed_i <- 111
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 35
```

@@

```{r gp-priors}
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
d.p = c(10, 1/0.1)
```

@@

```{r}
gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
gp_dat <- gp_predict(gp, x_grid)

@@

```{r}
plots <- summary_gp_mcmc(gp)
plots[[1]]
plots[[2]]
```

```{r}
# Summarize the GP model
tgp_dat <- 
    data.frame(  x = x_grid, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
```

@@

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 0.01, 40)
logr_prior <- function(x) dunif(x, -6, 6)
logtheta_prior <- function(x) dunif(x, -6, 6)
stdQ_prior <- function(x) dunif(x, 0.001, 100)
stdR_prior <- function(x) dunif(x, 0.001, 100)

@@

@@

plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

@@

```{r}
MaxT = 1000
matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
